Immutable Source-Mapped Abstract Syntax Tree:\ \, \\ A Design Pattern for Refactoring Engine APIs

### 2017-09-27 paper.tex
Promoted variants back to top-level section
Moved source-mapping solution to the solution section
Moved immutability argument to solution section
Lengthened italics solution text

The easiest way to map AST nodes to source locations is to store source
location information in the AST nodes themselves.  Again, consider the JDT AST
shown in Figure~\ref{fig:jdt-ast}, and note the numbers in square brackets.
The highlighted node is displayed as ``VariableDeclarationFragment [35, 8]'':
this AST node corresponds to the eight-character string beginning at offset 35
(\ttt{FIVE = 5}).  Its child node corresponding to the identifier \ttt{FIVE} is
attributed with the offset-length pair [35, 4], while the numeric literal
\ttt{5}'s source location is [42, 1].

@@

Mutable ASTs were used in the original Smalltalk Refactoring
Browser~\cite{roberts97refactoring}, as well as in
CRefactory~\cite{garrido05program} and Photran~\cite{sle2008}.  Many compilers
use mutable ASTs internally.  However, these tools have one thing in common:
the mutable AST is not accessible via any public API.  It is intended for
internal use only, by those tools' developers who should be better equipped to handle all of
the caveats.

@@

\textit{Use a parser for the language being refactored to construct an
\textup{abstract syntax tree}, or \textup{AST}.  Make the AST
\textup{immutable}---i.e., its structure cannot be changed---but provide
\textup{source mapping} information so that every node in the AST can be mapped
to a particular substring of the original source code.}

@@

\subsection{Variants}

>>>>

@@

@@

\textit{Use a parser for the language being refactored to construct an
\textup{abstract syntax tree}, or \textup{AST}.  Make the AST
\textup{immutable}---i.e., its structure cannot be changed---but provide
\textup{source mapping} information so that every node in the AST can be mapped
to a particular substring of the original source code. Treat the source code
as a string, and use the source code location information from the AST to
add, replace, and delete substrings.}

\subsection{Immutability}
As seen in Section 3.4, mutability introduces a host of problems to ASTs. Mutable ASTs were used in the original Smalltalk Refactoring
Browser~\cite{roberts97refactoring}, as well as in
CRefactory~\cite{garrido05program} and Photran~\cite{sle2008}.  Many compilers
use mutable ASTs internally.  However, these tools have one thing in common:
the mutable AST is not accessible via any public API.  It is intended for
internal use only, by those tools' developers who should be better equipped to handle all of
the caveats.

\subsection{Source Mapping}

As discussed in Section 3.3, it must be possible to map nodes in the AST to
locations in the source code, and vice versa. The easiest way to map AST nodes to source locations is to store source
location information in the AST nodes themselves.  Again, consider the JDT AST
shown in Figure~\ref{fig:jdt-ast}, and note the numbers in square brackets.
The highlighted node is displayed as ``VariableDeclarationFragment [35, 8]'':
this AST node corresponds to the eight-character string beginning at offset 35
(\ttt{FIVE = 5}).  Its child node corresponding to the identifier \ttt{FIVE} is
attributed with the offset-length pair [35, 4], while the numeric literal
\ttt{5}'s source location is [42, 1].

@@

\section{Variants}
